---
title: Fire science workshop materials 
author: |
  | Devan Allen McGranahan $*$  
  |
  | $*$ USDA-ARS
  |
date: |
  | Version date: `r format(Sys.time(), '%d %B %Y')`
output:
  tufte::tufte_handout:
    keep_tex: true
header-includes:
  - \renewcommand\familydefault{\sfdefault}
bibliography: FireScienceWorkshop.bib
link-citations: yes
csl: ijwf.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.show='hold', warning = FALSE, message = FALSE, echo=FALSE)
pacman::p_load(tint, tidyverse, knitr, pander)

```

# Summary 

Here's a citation [@mcgranahan2021]. 

## Findings

# References

# Headings

This style provides first and second-level headings (that is, `#` and `##`), demonstrated in the next section. You may get unexpected output if you try to use `###` and smaller headings.

`r newthought('In his later books')`^[[Beautiful Evidence](https://www.edwardtufte.com/tufte/books_be)], Tufte starts each section with a bit of vertical space, a non-indented paragraph, and sets the first few words of the sentence in small caps. To accomplish this using this style, call the `newthought()` function in **tufte** in an _inline R expression_ `` `r ` `` as demonstrated at the beginning of this paragraph.^[Note you should not assume **tufte** has been attached to your R session. You should either `library(tufte)` in your R Markdown document before you call `newthought()`, or use `tint::newthought()`.]

# Figures

## Margin Figures

Images and graphics play an integral role in Tufte's work. To place figures in the margin you can use the **knitr** chunk option `fig.margin = TRUE`. For example:

```{r fig-margin, fig.margin = TRUE, fig.cap = "MPG vs horsepower, colored by transmission.", fig.width=3.5, fig.height=3.5, cache=TRUE}
library(ggplot2)
mtcars2 <- mtcars
mtcars2$am <- factor(
  mtcars$am, labels = c('automatic', 'manual')
)
ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point() + geom_smooth() +
  theme(legend.position = 'bottom')
```

Note the use of the `fig.cap` chunk option to provide a figure caption. You can adjust the proportions of figures using the `fig.width` and `fig.height` chunk options. These are specified in inches, and will be automatically scaled down to fit within the handout margin.

## Arbitrary Margin Content

In fact, you can include anything in the margin using the **knitr** engine named `marginfigure`. Unlike R code chunks ```` ```{r} ````, you write a chunk starting with ```` ```{marginfigure} ```` instead, then put the content in the chunk. See an example on the right about the first fundamental theorem of calculus.

```{marginfigure}
We know from _the first fundamental theorem of calculus_ that for $x$ in $[a, b]$:
$$\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x).$$
```

For the sake of portability between LaTeX and HTML, you should keep the margin content as simple as possible (syntax-wise) in the `marginefigure` blocks. You may use simple Markdown syntax like `**bold**` and `_italic_` text, but please refrain from using footnotes, citations, or block-level elements (e.g. blockquotes and lists) there.

## Full Width Figures

You can arrange for figures to span across the entire page by using the chunk option `fig.fullwidth = TRUE`.

```{r fig-fullwidth, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "A full width figure.", warning=FALSE, cache=TRUE}
ggplot(diamonds, aes(carat, price)) + geom_smooth() +
  facet_grid(~ cut)
```

 Other chunk options related to figures can still be used, such as `fig.width`, `fig.cap`, `out.width`, and so on. For full width figures, usually `fig.width` is large and `fig.height` is small. In the above example, the plot size is $10 \times 2$.

## Main Column Figures

Besides margin and full width figures, you can of course also include figures constrained to the main column. This is the default type of figures in the LaTeX/HTML output.

```{r fig-main, fig.cap = "A figure in the main column.", cache=TRUE}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
```

# Sidenotes

One of the most prominent and distinctive features of this style is the extensive use of sidenotes. There is a wide margin to provide ample room for sidenotes and small figures. Any use of a footnote will automatically be converted to a sidenote. ^[This is a sidenote that was entered using a footnote.] 

If you'd like to place ancillary information in the margin without the sidenote mark (the superscript number), you can use the `margin_note()` function from **tufte** in an inline R expression. `r margin_note("This is a margin note.  Notice that there is no number preceding the note.")` This function does not process the text with Pandoc, so Markdown syntax will not work here. If you need to write anything in Markdown syntax, please use the `marginfigure` block described previously.

# References

References can be displayed as margin notes for HTML output. For example, we can cite a paper here [@mcgranahan2021]. 

# Tables

You can use the `kable()` function from the **knitr** package to format tables that integrate well with the rest of the Tufte handout style. The table captions are placed in the margin like figures in the HTML output.

```{r}
knitr::kable(
  mtcars[1:6, 1:6], caption = 'A subset of mtcars.'
)
```

# Block Quotes

We know from the Markdown syntax that paragraphs that start with `>` are converted to block quotes. If you want to add a right-aligned footer for the quote, you may use the function `quote_footer()` from **tufte** in an inline R expression. Here is an example:

> "If it weren't for my lawyer, I'd still be in prison. It went a lot faster with two people digging."
>
> `r tint::quote_footer('--- Joe Martin')`

Without using `quote_footer()`, it looks like this (the second line is just a normal paragraph):

> "Great people talk about ideas, average people talk about things, and small people talk about wine."
>
> --- Fran Lebowitz

# Some Notes on Tufte CSS

There are a few other things in Tufte CSS that we have not mentioned so far. If you prefer `r sans_serif('sans-serif fonts')`, use the function `sans_serif()` in **tufte**. For epigraphs, you may use a pair of underscores to make the paragraph italic in a block quote, e.g.

> _I can win an argument on any topic, against any opponent. People know this, and steer clear of me at parties. Often, as a sign of their great respect, they don't even invite me._
>
> `r quote_footer('--- Dave Barry')`

# References
